---
title: "HW2_WuJiaqi_119020436"
author: "WuJiaqi 119020436"
date: "10/26/2021"
output: html_document
---
# Q1: Work on the “dj.csv” index data which contains DJ index daily data.
# a. Subset your data to dates between 1985-01-01 and 2007-12-31.
```{r}
setwd("/Users/wujiaqi/Desktop/FIN3380/HW/HW2data")
dj = read.csv("dj.csv")
library(dplyr)
library(zoo)
library(tidyr)
dj = dj %>% filter(date > "1985-01-01", date < "2007-12-31")
head(dj)
tail(dj)
```

# b. Construct a 7 day RSI for your DJ data
```{r}
dj.rsi = dj %>%
  select(date,dji) %>%
  arrange(date) %>%
  mutate(date=as.Date(date),
         delta=dji-lag(dji,1))

dj.rsi = dj.rsi %>%
  mutate(up=ifelse(delta>0,1,0),
         down=ifelse(delta<0,1,0))

dj.rsi = dj.rsi %>%
  mutate(up.val=delta*up,
         down.val=-delta*down)  
head(dj.rsi)
```

```{r}
dj.rsi = dj.rsi[2:nrow(dj.rsi),] %>%
  mutate(up.first.avg=rollapply(up.val,
                                width=7,
                                align="right",
                                fill=NA,FUN=mean,na.rm=T),
         down.first.avg=rollapply(down.val,
                                  width=7,
                                  align="right",
                                  fill=NA,FUN=mean,na.rm=T))

dj.rsi$up.avg = dj.rsi$up.first.avg
dj.rsi$down.avg = dj.rsi$down.first.avg
for (t in 8:length(dj.rsi$up.avg)){
  dj.rsi$up.avg[t] = dj.rsi$up.avg[t-1]*6/7 + dj.rsi$up.val[t]*1/7
  dj.rsi$down.avg[t] = dj.rsi$down.avg[t-1]*6/7 + dj.rsi$down.val[t]*1/7
}

dj.rsi$rs = dj.rsi$up.avg/dj.rsi$down.avg
dj.rsi$rsi = 100-100/(1+dj.rsi$rs)
head(dj.rsi,10)
```

```{r}
plot(dj.rsi$rsi~dj.rsi$date,
     ylab = "RSI (7-Day Moving Avg)",
     xlab = "Date",
     ylim = c(0,100),
     type = "l",
     main = "DJ: RSI (7-Day Moving Avg)")
abline(h=c(30,70),col="red")  
par(new=T) 
plot(dj.rsi$dji~dj.rsi$date,
     xlab = "",
     ylab = "",
     yaxt = "n",  
     type = "l", col="lightblue")
axis(4) 
```

# c. How many days in total that the RSI is above or equal 70? How many days in total that the RSI is below or equal 30?
```{r}
cat("There are ",length(which(dj.rsi$rsi>=70))," days that RSI above or equal to 70")
cat("There are ",length(which(dj.rsi$rsi<=30))," days that RSI below or equal to 30")
```

# d. Check the performance of the following strategy: Buy the dji index at the close price when the rsi rises from below the oversold line (line 30), sell it at the close price on the following day. That is buy the stock if yesterday’s RSI < 30 but today’s RSI >= 30.

```{r}
dj.rsi = dj.rsi %>%
  arrange(date) %>%
  mutate(lag.rsi = lag(rsi,1)) %>%
  mutate(sellprice=lead(dji,1)) 
head(dj.rsi,10)
```

```{r}
dj.d = dj.rsi %>%
  select(date,dji,sellprice,rsi,lag.rsi) %>%
  filter(rsi>=30, lag.rsi<30)
head(dj.d)
```

```{r}
dj.d = dj.d %>%
  arrange(date) %>%
  mutate(ret=sellprice/dji-1) %>%
  mutate(cumret=cumprod(ret+1))
tail(dj.d)
cat("The cumret of this strategy is ",dj.d[nrow(dj.d),]$cumret)
```

### Q2: Work on crsp.sample.RData:
### a. Calculate your Equal-weighted index return over the entire periods, however, rather than reallocate monthly, reallocate your portfolio at the begining of each quarter, that is, the reallocation dates are the first trading date after 01-01, 04-01, 07-01, 10-01 each year. (Hint: use function as.yearqtr from the library(zoo)). Plot your index cumulative return on first quarter of 2012. Plot your cumulative index performance over the entire trading periods.
```{r}
load("/Users/wujiaqi/Desktop/FIN3380/HW/HW2data/crsp.sample.Rdata")
crsp.sample = crsp.sample %>% select(permno,date,ret)
tmp.summary = crsp.sample %>% group_by(permno) %>% summarise(n=n()) %>%   
  ungroup()
tmp.summary
```

```{r}
crsp.sample = spread(crsp.sample, key = permno, value = ret)
crsp.sample = gather(crsp.sample, key = permno, value = ret, -date)
tmp.summary = crsp.sample %>% group_by(permno) %>% summarise(n=n()) %>%
  ungroup()
tmp.summary
```

```{r}
crsp.sample = crsp.sample %>%
  mutate(ret = ifelse(is.na(ret),0,ret)) %>% 
  mutate(yyss = as.yearqtr(date)) %>%
  group_by(permno, yyss) %>%
  arrange(date) %>%
  mutate(cumret = cumprod(ret+1)) %>%
  ungroup()
head(crsp.sample,3)
```

```{r}
crsp.ew = crsp.sample %>%
  group_by(date) %>%
  summarise(ew = mean(cumret),
            yyss = first(yyss)) %>%
  ungroup()
head(crsp.ew)
```

```{r}
crsp.ew = crsp.ew %>%
  group_by(yyss) %>%
  arrange(date) %>%
  mutate(ret=ew/lag(ew,1)-1) %>% ungroup()
head(crsp.ew)
```

```{r}
crsp.ew = crsp.ew %>%
  mutate(ret = ifelse(is.na(ret),ew-1,ret))
crsp.ew = crsp.ew %>%
  arrange(date) %>%
  mutate(ew.all = cumprod(ret+1))
head(crsp.ew)
```

```{r}
crsp.ew.2012Q1 = crsp.ew %>% filter(yyss == "2012 Q1")
plot(crsp.ew.2012Q1$ew.all~crsp.ew.2012Q1$date,
     type = "l",
     xlab = "date",
     ylab = "ew.id",
     main = "Equal Weighted Index (2012 Q1)")
```

```{r}
plot(crsp.ew$ew.all~crsp.ew$date,
     type = "l",
     xlab = "date",
     ylab = "ew.id",
     main = "Equal Weighted Index (entire period)")
```

### b. Repeat a. to construct Value-weighted index with quarterly reallocation
```{r}
load("/Users/wujiaqi/Desktop/FIN3380/HW/HW2data/crsp.sample.Rdata")
crsp.sample = crsp.sample %>%
  mutate(me = abs(prc)*shrout) %>%
  select(permno,date,ret,me) %>%     
  mutate(permno = as.character(permno))
head(crsp.sample)
```

```{r}
crsp.ret = spread(crsp.sample[,c("permno","date","ret")],permno,ret)
crsp.ret = gather(crsp.ret,key=permno,value=ret,-date)
crsp.ret = crsp.ret %>% mutate(ret=ifelse(is.na(ret),0,ret))
crsp.me = crsp.sample %>% select(permno,date,me)
crsp.sample = crsp.ret %>% left_join(crsp.me,by=c("permno","date"))
head(crsp.sample)
```

```{r}
crsp.vw = crsp.sample %>%
  group_by(permno) %>%
  arrange(date) %>%
  mutate(lag.me = lag(me,1)) %>% ungroup()
head(crsp.vw,20)
```

```{r}
crsp.vw = crsp.vw %>%
  mutate(yyss=as.yearqtr(date)) %>%
  group_by(permno,yyss) %>%
  arrange(date) %>%
  mutate(lag.me=first(lag.me)) %>% ungroup()  
tail(crsp.vw,20)
```

```{r}
crsp.vw = crsp.vw %>% filter(yyss!=first(yyss))
crsp.vw = crsp.vw %>%
  group_by(date) %>%
  mutate(wgt=lag.me/sum(lag.me,na.rm=T)) %>%
  ungroup()
head(crsp.vw,10)
```

```{r}
crsp.vw = crsp.vw %>%
  group_by(yyss,permno) %>%
  arrange(date) %>%
  mutate(cumret=cumprod(ret+1)) %>% 
  ungroup() %>% 
  arrange(date,permno)
head(crsp.vw,20)
```

```{r}
crsp.vw = crsp.vw %>%
  mutate(vw=wgt*cumret)

crsp.vw = crsp.vw %>%
  group_by(date) %>%
  summarize(vw=sum(vw),
            yyss=first(yyss)) %>% 
  ungroup()

crsp.vw = crsp.vw %>%
  group_by(yyss) %>%
  arrange(date) %>%
  mutate(ret=vw/lag(vw,1)-1) %>% 
  ungroup()
tail(crsp.vw)
```

```{r}
crsp.vw = crsp.vw %>% mutate(ret=ifelse(is.na(ret),vw-1,ret))
crsp.vw = crsp.vw %>%
  arrange(date) %>%
  mutate(vw.all=cumprod(ret+1))

crsp.vw.2012Q1 = crsp.vw %>% filter(yyss=="2012 Q1")
yrange = range(crsp.vw.2012Q1$vw.all)
plot(crsp.vw.2012Q1$vw.all~as.Date(crsp.vw.2012Q1$date),
     type="l", xlab="date",ylab="index",
     main="Value Weighted Index 2012 Q1",
     ylim=yrange)
```

```{r}
plot(crsp.vw$vw.all~as.Date(crsp.vw$date),
     type="l", xlab="date", ylab="vw.id",
     ylim=range(crsp.vw$vw.all),
     main="Value Weighted Index (entire period)")
```

### Q3: Work on data “crsp.sample2.RData”
### a. When is the first date that each stock start to have return data?
```{r}
load("/Users/wujiaqi/Desktop/FIN3380/HW/HW2data/crsp.sample2.Rdata")
crsp.sample2 = crsp.sample2 %>%
  group_by(permno) %>%
  arrange(date)  %>%
  mutate(fd = first(date)) %>%
  ungroup()
head(crsp.sample2,10)
```

```{r}
stock.id = unique(crsp.sample2$permno)
stock.id
```

```{r}
for (i in 1:10){
  firstdate = as.character(crsp.sample2[which(crsp.sample2$permno==stock.id[i]),][1,]$fd)
  q = paste0("The first date that stock ",stock.id[i]," starts to have return is ", firstdate)
  print(q)
}
```

### b. Assume the first date you achieve from a. is the launch date(IPO date) of each stock, build your monthly reallocation equal-weighted portfolio index taking into account of launch date of a stock. For example, on “2011-01-04”, if only 9 out of 10 stocks launched, your index should only contain 9 stocks weighted average rather than 10 stocks. Plot the cumulative return of your index in 2013-01 and 2011-01.
```{r}
crsp.sample2.a = crsp.sample2 %>% 
  filter(date<"2012-01-03") %>%
  select(permno,date,ret)
tmp.summary = crsp.sample2.a %>% group_by(permno) %>% summarise(n=n()) %>%   
  ungroup()
tmp.summary
```

```{r}
crsp.sample2.a = crsp.sample2.a %>%
  mutate(ret = ifelse(is.na(ret),0,ret)) %>%  
  mutate(date = as.character(date)) %>%
  mutate(yymm = substr(date,1,7)) %>%
  group_by(permno, yymm) %>%
  arrange(date) %>%
  mutate(cumret = cumprod(ret+1)) %>%
  ungroup()
head(crsp.sample2.a,3)
```

```{r}
crsp2a.ew = crsp.sample2.a %>%
  group_by(date) %>%
  summarise(ew = mean(cumret),
            yymm = first(yymm)) %>%
  ungroup()

crsp2a.ew = crsp2a.ew %>%
  group_by(yymm) %>%
  arrange(date) %>%
  mutate(ret=ew/lag(ew,1)-1) %>% 
  ungroup()

crsp2a.ew = crsp2a.ew %>%
  mutate(ret = ifelse(is.na(ret),ew-1,ret))
crsp2a.ew = crsp2a.ew %>%
  arrange(date) %>%
  mutate(ew.all = cumprod(ret+1))
tail(crsp2a.ew)
```

```{r}
crsp2a.ew.2011.01 = crsp2a.ew %>% filter(yymm == "2011-01")
plot(crsp2a.ew.2011.01$ew.all~as.Date(crsp2a.ew.2011.01$date),
     type = "l",
     xlab = "date",
     ylab = "ew.id",
     main = "Equal Weighted Index (2011-01)")
```

```{r}
crsp.sample2.b = crsp.sample2 %>% 
  filter(date>="2012-01-03") %>%
  select(permno,date,ret)
tmp.summary = crsp.sample2.b %>% group_by(permno) %>% summarise(n=n()) %>%   
  ungroup()
tmp.summary
```

```{r}
crsp.sample2.b = spread(crsp.sample2.b, key = permno, value = ret)
crsp.sample2.b = gather(crsp.sample2.b, key = permno, value = ret, -date)
tmp.summary = crsp.sample2.b %>% group_by(permno) %>% summarise(n=n()) %>%
  ungroup()
tmp.summary
```

```{r}
crsp.sample2.b = crsp.sample2.b %>%
  mutate(ret = ifelse(is.na(ret),0,ret)) %>%  
  mutate(date = as.character(date)) %>%
  mutate(yymm = substr(date,1,7)) %>%
  group_by(permno, yymm) %>%
  arrange(date) %>%
  mutate(cumret = cumprod(ret+1)) %>%
  ungroup()
head(crsp.sample2.b,3)
```

```{r}
crsp2b.ew = crsp.sample2.b %>%
  group_by(date) %>%
  summarise(ew = mean(cumret),
            yymm = first(yymm)) %>%
  ungroup()

crsp2b.ew = crsp2b.ew %>%
  group_by(yymm) %>%
  arrange(date) %>%
  mutate(ret=ew/lag(ew,1)-1) %>% ungroup()

crsp2b.ew = crsp2b.ew %>%
  mutate(ret = ifelse(is.na(ret),ew-1,ret))
head(crsp2b.ew)
```

```{r}
crsp2.ew = rbind(crsp2a.ew[,1:4],crsp2b.ew)
crsp2.ew

crsp2.ew = crsp2.ew %>%
  arrange(date) %>%
  mutate(ew.all = cumprod(ret+1))
head(crsp2.ew)
tail(crsp2.ew)
```

```{r}
crsp2.ew.2013.01 = crsp2.ew %>% filter(yymm == "2013-01")
tail(crsp2.ew.2013.01)
plot(crsp2.ew.2013.01$ew.all~as.Date(crsp2.ew.2013.01$date),
     type = "l",
     xlab = "date",
     ylab = "ew.id",
     main = "Equal Weighted Index (2013-01)")

```

### Q4 The risk-parity investment strategy is to allocate your investment according to the inverse of each asset’s riskness. This allocation strategy is shown to be better than the value-weighted allocation. Consider a portfolio contains 6 stocks “AAPL”,“MSFT”,“IBM”,“BABA”,“GE”,“KO” as shown in class. Reallocate your portfolio monthly. At the begining of month t, the weight of asset i is 1/sd(i,t−1)Σ6j=1[1/sd(j,t−1)] where sd(i,t−1) is the monthly standard deviation of the stock i at month t-1.
### a. What are the weights of each asset in “2012-03”?
```{r}
load("/Users/wujiaqi/Desktop/FIN3380/HW/HW2data/mystock.Rdata")
mystock = mystock %>%
  group_by(permno) %>%
  arrange(date) %>%
  mutate(ret=prc/lag(prc,1)-1) %>%
  ungroup()
head(mystock)
tmp.summary = mystock %>% group_by(permno) %>% summarise(n=n()) %>%
  ungroup()
tmp.summary
```

```{r}
mystock.ret = spread(mystock[,c(1,2,5)], key = permno, value = ret)
mystock.ret = gather(mystock.ret, key = permno, value = ret, -date)
mystock.ret = mystock.ret %>% 
  mutate(permno = as.numeric(permno))
mystock.other = mystock[,c(1,2,3,4)]

mystock = mystock.ret %>%
  left_join(mystock.other, by = c("permno","date")) %>%
  arrange(date)
head(mystock,12)
```

```{r}
mystock = mystock %>%
  mutate(prc = ifelse(is.na(prc),0,prc),
         ret = ifelse(is.na(ret),0,ret),
         yymm = substr(date,1,7))

mystock = mystock %>%
  group_by(permno,yymm) %>%
  mutate(std = sd(ret)) %>% 
  ungroup()

mystock = mystock %>% 
  group_by(permno) %>%
  mutate(lag.std = lag(std,1)) %>%
  ungroup()
head(mystock,20)
```

```{r}
mystock = mystock %>%
  filter(yymm > "2010-01") %>%
  group_by(permno,yymm) %>%
  arrange(date) %>%
  mutate(lag.std = first(lag.std)^(-1),
         lag.std = ifelse(!is.finite(lag.std),0,lag.std))

mystock = mystock %>%
  group_by(date) %>%
  mutate(wgt=lag.std/sum(lag.std)) %>%
  group_by(permno,yymm) %>%
  arrange(date) %>%
  mutate(cumret=cumprod(ret+1)) %>%
  ungroup()
head(mystock)
```

```{r}
mystock.201203 = mystock %>%
  filter(yymm == "2012-03") 
head(mystock.201203)
```

### b. Plot your index performance in year 2013 together with the index performance built based on Equal Weighted and Value-Weighted.

```{r}
# RPI
mystock.rpi = mystock %>%
  mutate(rpi = wgt*cumret)

mystock.rpi = mystock.rpi %>%
  group_by(date) %>%
  summarize(rpi=sum(rpi),
            yymm=first(yymm)) %>% 
  ungroup()
tail(mystock.rpi)
```

```{r}
mystock.rpi = mystock.rpi %>%
  group_by(yymm) %>%
  arrange(date) %>%
  mutate(ret=rpi/lag(rpi,1)-1) %>% 
  ungroup()
tail(mystock.rpi)
```

```{r}
mystock.rpi = mystock.rpi %>% 
  mutate(ret=ifelse(is.na(ret),rpi-1,ret)) %>%
  arrange(date) %>%
  mutate(rpi.all=cumprod(ret+1))

mystock.rpi.2013 = mystock.rpi %>% filter(yymm >= "2013-01", yymm <= "2013-12")
```

```{r}
# EW
load("/Users/wujiaqi/Desktop/FIN3380/HW/HW2data/mystock.Rdata")
mystock.ew = mystock %>%
  group_by(permno) %>%
  arrange(date) %>%
  mutate(ret=prc/lag(prc,1)-1) %>%
  mutate(me = abs(prc)*shrout) %>%   
  mutate(permno = as.character(permno)) %>%
  select(permno,date,ret) %>%    
  ungroup()
head(mystock.ew,10)
```

```{r}
mystock.ew = mystock.ew %>%
  mutate(yymm = substr(date,1,7)) %>%
  group_by(permno,yymm) %>%
  arrange(date) %>%
  mutate(cumret=cumprod(ret+1)) %>%
  ungroup()
  
mystock.ew = mystock.ew %>%
  filter(yymm >= "2010-01") %>%
  group_by(date) %>%
  summarise(ew = mean(cumret),
            yymm = first(yymm)) %>%
  ungroup()
head(mystock.ew)
```

```{r}
mystock.ew = mystock.ew %>%
  group_by(yymm) %>%
  arrange(date) %>%
  mutate(ret=ew/lag(ew,1)-1) %>% ungroup()
head(mystock.ew)
mystock.ew = mystock.ew %>%
  mutate(ret = ifelse(is.na(ret),ew-1,ret)) %>%
  filter(ret != 0)

mystock.ew = mystock.ew %>%
  arrange(date) %>%
  mutate(ew.all = cumprod(ret+1))
tail(mystock.ew)

mystock.ew.2013 = mystock.ew %>% filter(yymm >= "2013-01", yymm <= "2013-12")
```

```{r}
# VW
load("/Users/wujiaqi/Desktop/FIN3380/HW/HW2data/mystock.Rdata")
mystock = mystock %>%
  mutate(me = abs(prc)*shrout) %>%   
  mutate(permno = as.character(permno)) %>%
  group_by(permno) %>%
  arrange(date) %>%
  mutate(ret=prc/lag(prc,1)-1) %>%
  select(permno,date,ret,me) %>%    
  ungroup()
head(mystock,10)
```

```{r}
mystock.vw = spread(mystock[,c("permno","date","ret")],permno,ret)
mystock.vw = gather(mystock,key=permno,value=ret,-date)
mystock.vw = mystock %>% mutate(ret=ifelse(is.na(ret),0,ret))
head(mystock.vw)
```

```{r}
mystock.vw = mystock.vw %>%
  group_by(permno) %>%
  arrange(date) %>%
  mutate(lag.me = lag(me,1)) %>% ungroup()

mystock.vw = mystock.vw %>%
  mutate(yymm=substr(date,1,7)) %>%
  group_by(permno,yymm) %>%
  arrange(date) %>%
  mutate(lag.me=first(lag.me)) %>% ungroup()  

mystock.vw = mystock.vw %>% filter(yymm!=first(yymm))
head(mystock.vw)
```

```{r}
mystock.vw = mystock.vw %>%
  group_by(date) %>%
  mutate(wgt=lag.me/sum(lag.me,na.rm=T)) %>%
  ungroup()

mystock.vw = mystock.vw %>%
  group_by(yymm,permno) %>%
  arrange(date) %>%
  mutate(cumret=cumprod(ret+1)) %>% ungroup()
mystock.vw = mystock.vw %>% arrange(date,permno)
head(mystock.vw)
```

```{r}
mystock.vw = mystock.vw %>%
  mutate(vw=wgt*cumret)

mystock.vw = mystock.vw %>%
  group_by(date) %>%
  summarize(vw=sum(vw),
            yymm=first(yymm)) %>% ungroup()

mystock.vw = mystock.vw %>%
  group_by(yymm) %>%
  arrange(date) %>%
  mutate(ret=vw/lag(vw,1)-1) %>% ungroup()
head(mystock.vw)
```

```{r}
mystock.vw = mystock.vw %>% 
  mutate(ret=ifelse(is.na(ret),vw-1,ret)) %>%
  filter(ret != 0)

mystock.vw = mystock.vw %>%
  arrange(date) %>%
  mutate(vw.all=cumprod(ret+1))
tail(mystock.vw)

mystock.vw.2013 = mystock.vw %>% filter(yymm >= "2013-01", yymm <= "2013-12")
```

```{r}
plot(mystock.rpi.2013$rpi.all~as.Date(mystock.rpi.2013$date),
     type="l", xlab="date", ylab="return", 
     ylim=range(mystock.rpi.2013$rpi.all,mystock.ew.2013$ew.all,mystock.vw.2013$vw.all),
     main="Index Performance")
lines(mystock.ew.2013$ew.all~as.Date(mystock.ew.2013$date),
      col="red",
      type="l")
lines(mystock.vw.2013$vw.all~as.Date(mystock.vw.2013$date),
     col="blue",
     type="l")
legend("topleft",                              
       c("RPI","EW","VW"),
       lty = c(1,1,1),                          
       col = c("black","red","blue"))
```













